### 2019-02-20 paper2018/paper.tex
background part added

\documentclass{article}
\usepackage{todonotes}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

\usepackage[backend=bibtex,citestyle=authoryear,style=alphabetic,natbib]{biblatex}
\addbibresource{PaperTools/bibtex/jp.bib}
\addbibresource{phorae.bib}
\addbibresource{IWCSbib.bib}

\usepackage{stmaryrd}
\usepackage{enumitem}
\usepackage{newunicodechar}
\input{PaperTools/latex/newunicodedefs}

\title{FraCaS: 80 percent complete}

% - Two-phase system (Dyn semantics from theorem proving)
% - Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
% - Multiple readings
% - Threshold-based interpretation of (some) adjectives + Linear arithmetic proofs
% - New handling of prepositions and adverbs
% - Definites handled by adding an assumption at the top-level. (instead as inline existential) -- Enabled by dyn. semantics
% - Plurals/Quantifiers improved
% - Genders are properties of nouns
% - Handling of comparatives (with threshold BUT not those of adjectives), relying largely on the dyn. semantic system

\newlist{lingex}{enumerate}{3}
\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}

\begin{document}

\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether an NL hypothesis H follows from an NL premise(s) P. NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLI. The latter approaches, following the general trend in NLP, have been dominating NLI and indeed a number of impressive results have been produced \citet{kim:2018,radford:2018,liu:2019}.\footnote{These are the three systems with the best results on SNLI in increasing order.} State-of-the-art DL systems achieve an accuracy of around 0.9 when tested on suitable datasets. However, the datasets that are used are assuming a definition of inference that can be thought to be `looser' or less precise compared to the definition assumed in platforms that used to test logical approaches. Consider the following two examples, one from the SNLI dataset, predominatly used to test DL approaches and which has a state-of-the-art performace of 0.911 \citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par excellence} to test logical approaches.

\begin{lingex}
\item An SNLI example \\
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction}
\end{lingex}

\begin{lingex}
\item An example from FraCaS\\

\textbf{P1}] A Scandinavian won the Nobel Prize.\\
\textbf{P2}	Every Swede is  Scandinavian.\\
\textbf{Q}  Did a Swede win the Nobel prize? \\
\textbf{H} A Swede won the Nobel prize.\\
\textbf{Label: Yes} [FraCaS 173]
\end{lingex}

In the SNLI example, one has to make a number of non-trivial assumptions in order to arrive at a contradiction:  a)  the two situations described have to be taken to refer to  the same situation and, thus, decide  whether one description contradicts the other, b) the indefinite article in the premise has to be identified with the indefinite article in the hypothesis, i.e.  \textit{man} contradicting \textit{woman} (a person cannot be a man and a woman at the same time). This is part of the reasoning humans perform, but it is not the only one. More precise, logical reasoning can be also performed. Such an example is the FraCaS example, as well as very fine-grained cases like the Antecedent Contained Deletion  ellipsis example below:

\begin{lingex}

\item An example from FraCaS involving fine-grained reasoning\\
\textbf{P1}	Bill spoke to everyone that John did.	\\	
\textbf{P2}	John spoke to Mary.	\\
\textbf{Q} 	Did Bill speak to Mary?\\
\textbf{H} 	Bill spoke to Mary.\\
\textbf{Label:	UNK} [FraCaS 065] 	\end{lingex}

For this reason, and despite the dominance of DL approaches in pretty much all NLP tasks, logical approaches have been continuing to be developed and evaluated on datasets like the FraCaS test suite and the SICK dataset \cite{marelli:2014}.  \citet{bernardy:2017}  define  a correspondence between abstract syntax parse trees of the FraCas examples, parsed using the Grammatical Framework (GF, \citet{Ranta:GF}), and modern type-theoretic semantics that are output in the Coq proof assistant (the FraCoq system).  The accuracy is 0.85 for 5 sections of the FraCaS test suite. The \textsc{Langpro} system presented by \citet{abzianidze:2017} is based on a Natural Logic tableau theorem prover. It achieves an accuracy of .82 on the SICK dataset.

In this paper, we concentrate on this sort of fine-grained, logical reasoning. In particular, we present the first system that deals with the sections on ellipsis and anaphora in the FraCaS test suite and has both the best coverage, as well as accuracy compared to other logical approaches. More specifically, the contributions of the paper are:

\begin{itemize}
 \item  A two-phase system (rather than one-phase as in  \citet{bernardy_type_2017}): a) using Haskell as a converter between Grammatical Framework trees and Coq types, b) using various type combinators to reach the desired NL semantics. 
  \item  The first run of the Anaphora and Ellipsis sections of the FraCaS (based on work by \citet{bernardy_jolli})
 \item  More coverage than any previous logical approach
  \item  State-of-the-art results in terms of accuracy compared to all previous approaches, for every section of the FraCaS
  \item  Handing of multiple readings
  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
  \item  New handling of prepositions and adverbs compared to the earlier system in  \citet{bernardy_type_2017}, as well as new handling of definitions, enabled by the dynamic semantics system. 
  \item  Improvement over the previous tre
atment of Plurals/Quantifiers
 \item  Big improvement on the comparatives section  enabled by handling  comparatives using a threshold (BUT one different to the one used for adjectives). Again, the dynamic semantics system is crucial here.

\end{itemize}
\section{Background: GF, Monadic Dynamic Semantics, Coq}

\section{GF}

\subsection{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete s usyntax of varionatural languages. These mappings can be inverted by GF, thus offering
parsing from natural text into abstract syntax. However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity. \textbf{Do you want to further comment here?}

\subsection{Coq}
Coq is an  interactive theorem prover (proof assistant)  based on the calculus of inductive
constructions (CiC), i.e.  a lambda calculus with dependent
types. Coq is arguably one of the leading proof assistants. It has
famously been instrumental in formalising a proof of the four-color
theorem \citep{Gonthier:2008}, a proof of the odd order theorem \citep{Gonthier:2013}, as well as developing CompCert, a formally verified compiler for C \citep{Leroy}. Coq is a very powerful reasoning engine that makes it fit for the task of NLI, when the latter is formalized as theorem proving task. Its underlying higher order logic is furthermore very expressive to express fine-grained logical semantics. Furthermore, Coq supports a number of constructions that we will be using notably dependent typing and subtyping. These features are used in order to take care of specific NL phenomena (e.g. subtyping is used in defining different adjectival inference classes).

\subsection{Dynamic Monadic Semantics}

\section{Overview of the system}

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program. It implements a dynamic
  semantics. It comprises the bigger part of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. This includes for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}
This architecture is represented schematically in Figure \ref{fig:overview}.

\begin{figure}
  \centering
{
  \tiny
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
                    semithick]
  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]

  \node[state]         (txt)                {text};
  \node[state]         (ast) [right of=txt] {AST};
  \node[state]         (pp)  [right of=ast] {Probabilistic program};
  \node[state]         (res) [right of=pp] {result in $(0,1)$};

  \path (txt) edge              node (GF)  {GF}                  (ast)
        (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
        (pp) edge               node (Coq) {Coq} (res);

  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};

  \path
  (Syntax) edge (GF)
  (Sem) edge (Hask)
  (Hyper) edge (Coq);
\end{tikzpicture}
}
\caption{Phases in our system. ($\ast$) At the level of Coq, we handle
  the details of the adverbial (veridicality properties) and
  adjectival semantics (division into subsective, extentional,
  non-committal, etc. categories.) }
  \label{fig:overview}
\end{figure}

All underlying systems (GF, Haskell, Coq) are based on lambda calculi
with types. We ensure that each translation preserve typing, locally:
\begin{enumerate}
\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
\item Every GF syntactic construction function $f : X$ is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
\end{enumerate}
Because all system embed the simply-typed lambda calculus, ensuring
type-preservation locally means that types are preserved globally.
Therefor, we are certain that every GF syntax tree can be mapped to
Haskell, and eventuall Coq, without error.

The dynamic semantics follows a monadic structure, as pioneered by
\citet{Shan:2002}. There are two kinds of effects carried by the
monad.  The first one comprises a series of updates and queries of
stateful elements.  There is one piece of updateable state for every
element which can be referred to by anaphoric expressions. These can
be the usual ones (like NPs), but also less usual ones (like 2-place
verbs, or a quantity --- which we illustrate below). The other kind of
effects is \emph{non-determinism}. We use non-determinism to model the
property that linguistic expressions can have several
interpretations. The monadic structure allows to locally express that
a given expression has several meanings; the monadic bind ensures that
all combinations of meanings are considered at the top-level,
combinatorially. This dynamic semantics allows us to model many
phenomena in a precise way.

\paragraph{anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples such as the following:
\begin{lingex}
\item[FraCaS 122]
P1	Every committee has a chairman.\\
P2	He is appointed its members.\\
H 	Every committee has a chairman appointed by members of the committee. 
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase. Such
ellptic expressions are handled in the same way as anaphora.
\begin{lingex}
\item[FraCas 173]
  P1	Bill spoke to everyone that John did [elliptic V2].\\
  P2	John spoke to Mary.\\
  Q 	Did Bill speak to Mary?\\
  H 	Bill spoke to Mary.
\end{lingex}
For example, in (FraCaS 173, P1), the argument of ``did'' is
explicitly marked as an elliptic V2, which we resolve to ``speak'' in
that context.
\paragraph{definites} A naive way to handle definites is an
existential (or sigma). However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

Instead, our take is that definites should be treated as an anaphoric
expression with an implicit referrent. That is, if the referent is not
found in the discourse, then it will be forcibly introduced, using an
existential type, \emph{at the top-level} of the expression. To be
able to do this, we record all definites without referent, using
another portion of the environment (using a monadic effect). For the
above example, we obtain the desired interpretation:
$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions. Finally, ``more'' is intepreted as ``at least
<elliptic quantity>+1''. This treatment, which we illustrated here on
an example, is systematic in our implementation.

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$). Formalization aside, this idea is reminiscent of degree-based approaches to gradable adjectives like \citet{Cresswell:1976,Heim:1985,Kennedy:2007}.   Additionally,
adjectival phrases, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable. Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

\paragraph{Adverbs}
Another point, of minor theoretical importance but major practical
one, is our handling of adverbial phrases. We interpret adverbs (and
in general all adverbial and prepositional phrases) as VP-modifiers:
$Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
to verb-phrases heavily complicates the Coq proofs, because such
phrases can contain quantifiers. Therefore, we instead move the
adverbs, so that they apply to (atomic) verbs only. Proofs can then be
simplify accordingly. \todo{theoretical justification in the
  literature?}

\section{Results and evaluation}

We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
total of 259 cases. We excluded only section 7, ``temporal reference''. The reason for
doing so is that, in our view, it contains too many examples which
require \textit{ad-hoc} treatment, and thus makes little sense to
include without complementing it with a more thorough suite which
captures a more complete landscape of the phenomena that section 7
touches.

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by Bill MacCartney\todo{cite} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labelling (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of the following cases, which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.

\begin{center}
\begin{tabular}{ccc}
 test case & new class & comment \\ \hline
    005 & UNK & missing hypothesis: there are italian tenors \\
   056 &  Yes & Already identified as such by MacCartney \\
  069 & Unk & Mary could have used someone else's workstation \\
  119 & Unk & \textit{ibid.} \\
  181 & Yes & for the same reason as 180 \\
  226 & Yes &
\end{tabular}
\end{center}

Our system classifies a case as YES if a proof can be constructed from
the premises to the hypothesis, NO if a proof of the negated
hypothesis can be constructed and UNK otherwise. Because we work with
a non-decidable logic, one cannot \emph{in general} conclude
decisively that no proof exists. Thus, we consider here that no proof
exists if it cannot be constructed with reasonable effort. In
particular, we test at the minimum that the automatic proof search
built in Coq does not succeed before classifying a problem as UNK.

%  ------ Section 1
% complete: 74/74   []
% correct: 71/74   [13,14,46]
% score: 0.9594594594594594
%  ------ Section 2
% complete: 33/33   []
% correct: 27/33   [84,91,94,95,109,111]
% score: 0.8181818181818182
%  ------ Section 3
% complete: 27/28   [137]
% correct: 24/27   [124,126,127]
% score: 0.8571428571428571
%  ------ Section 4
% complete: 47/52   [171,172,191,193,195]
% correct: 45/47   [170,188]
% score: 0.8653846153846154
%  ------ Section 5
% complete: 20/20   []
% correct: 19/20   [216]
% score: 0.95
%  ------ Section 6
% complete: 29/31   [244,245]
% correct: 27/29   [242,250]
% score: 0.8709677419354839
%  ------ Section 7
% complete: 0/67   [251,252,253,254,255,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,277,278,279,280,281,282,283,284,285,286,287,288,289,290,292,293,294,295,296,297,298,299,300,301,302,303,304,306,307,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325]
% correct: 0/0   []
% score: 0.0
%  ------ Section 8
% complete: 8/8   []
% correct: 6/8   [326,333]
% score: 0.75
%  ------ Section 9
% complete: 13/13   []
% correct: 12/13   [346]
% score: 0.9230769230769231
%  ------ Section 0
% complete: 251/259   [137,171,172,191,193,195,244,245]
% correct: 231/251   [13,14,46,84,91,94,95,109,111,124,126,127,170,188,216,242,250,326,333,346]
% score: 0.8918918918918919

\providecommand\forcecenter{\multicolumn{1}{c}}
\begin{table}
  \centering
\begin{tabular}{rlrrrrrr}
  & Section      & {\kern -1em} \# cases & Ours     & FraCoq & MINE & Nut  & Langpro  \\ \hline
1 & Quantifiers  & 75          & .96 (74) & .96    & .77  & .53  & .93 (44) \\
2 & Plurals      & 33          & .82      & .76    & .67  & .52  & .73 (24) \\
3 & Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
4 & Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
5 & Adjectives   & 22          & .95 (20) & .95    & .68  & .32  & .73 (12) \\
6 & Comparatives & 31          & .87      & .56    & .48  & .45  &  -       \\
7 & Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
8 & Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
9 & Attitudes    & 13          & .92      & .85    & .77  & .46  & .92 (9)  \\ \hline
  & Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
  &              &             & (259)    & (174)  & (174)& (174)& (89)
  \end{tabular}
  \caption{Overview of the accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases shown in the
    third column, the actual number of test cases attempted is shown in
    parentheses.  ``FraCoq'' refers to the work of
    \citet{bernardy_type_2017}. ``MINE" refers to the approach of
    \citet{Mineshima:2015}, ``NUT" to the CCG system that utilizes the
    first-order automated theorem prover \textit{nutcracker} described
    by \citet{bos:2008}, and ``Langpro" to the system presented by
    \citet{Abzianidze:2015}. A dash indicates that no attempt was made
    for a given section. }
  \label{tbl:results}
\end{table}

Table \ref{tbl:results} shows a considerable improvement over earlier
approaches in terms of coverage, with three more sections covered over
previous state of the art. We thus cover 259 out of 337 cases (77\%),
compared to at most 174 cases (52\%) in previous work. Additionally,
our system performs generally the best in terms of accuracy. In
particular, section 6 sees a large improvement in accuracy, which we
attribute to our new analysis of comparatives, enabled by our dynamic
semantics.

\paragraph{error analysis}
Our system fails to classify correctly 28 cases out of 259.  We give
here a summary of the missing features which are responsible for the
biggest number of failures. The one biggest source of error is an
incomplete handling of group readings. (FraCaS
013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
are cases where a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases are where a definite plural does
not have a universal reading (091,094,095) --- these require an
analysis at the pragmatic level, which our system lacks. Additionally,
neither measure phrases (242) nor attributive comparatives (244,245)
are handled.

\section{Future work}
\section{Conclusion}

\printbibliography

\end{document}